---
title: "Strandings project"
author: "Ellen Coombs"
date: "24 July 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Does climate influence cetacean distributions across the UK? 

This is an R markdown document for the cetacean strandings project. It includes: 

- The main questions that are being asked 
- My attempt to begin asking them
- An idea of the code I have used
- An idea of the plots I would produce 


##Questions 

##A. Cetacean distributions across the UK through time 
###i. Species richness - map/plot different species 
###ii. Abundance - histogram of species abundance (spatially as well) 
###iii. Species changes 

##B. Link to sea surface temperature (SST)

##C. Interesting side questions 
###i. Current seasonal changes in cetacean distributions 
###ii. Cross-validation of data: live survey data, bycatch data and whaling data 

##D. The 'boring' questions 
###i.Sampling effort - county data - need to work out how to get this on the hex plots 
###ii. Sampling bias - Basic plot (specific species/all species for that year)


```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
