---
title: "Strandings_Project_SST_Rmd"
author: "Ellen Coombs"
date: "07 August 2017"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

I have selected 14 locations around the UK and Ireland. Locations were selected based on whether they were the most 'northerly', "easterly" etc for the country, or whether they offered a general geographical location e.g. for the NW and were at "grid point nearest to [chosen geographical location]"

The mean of all of these locations was taken. The following is code for extracting data for all 14 locations, as well as the mean UK SST (degrees celsius), for the time period 1912-2016. 

Data used in HadISST1 monthly sea surface temperature (1870 - present)
Files are netCDF (.nc format) and were read in using the ncdf package

A lot of the code was used to test the data were correct and is put in here as comments 

##Working with HadISST1 data 
```{r setup, include=FALSE}

library(chron)
library(RColorBrewer)
library(lattice)
library(ncdf4)
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)

#Reading in .nc data - this is a huge dataset of 111,000,000 datapoints 
ncin <- nc_open("HadISST_sst.nc")

#Having a look at what's in the files 
attributes(ncin)$names
#[1] "filename"    "writable"    "id"          "safemode"    "format"      "is_GMT"     
#[7] "groups"      "fqgn2Rindex" "ndims"       "natts"       "dim"         "unlimdimid" 
#[13] "nvars"       "var"        

#Variables 
attributes(ncin$var)$names
#[1] "time_bnds" "sst" 

#Taking a look at the data 
ncin_data <- ncvar_get(ncin, attributes(ncin$var)$names[1])
print(ncin_data)

#Global attributes 
# get global attributes
title <- ncatt_get(ncin,0,"title")
institution <- ncatt_get(ncin,0,"institution")
datasource <- ncatt_get(ncin,0,"source")
references <- ncatt_get(ncin,0,"references")
history <- ncatt_get(ncin,0,"history")
Conventions <- ncatt_get(ncin,0,"Conventions")

# get longitude and latitude
lon <- ncvar_get(ncin,"longitude")
nlon <- dim(lon)
head(lon)
summary(lon)

lat <- ncvar_get(ncin,"latitude")
nlat <- dim(lat)
head(lat)
summary(lat)

print(c(nlon,nlat))

#Time variable from the dataset 
# get time

#Time since when? 
ncin$dim$time$units
#Looking at the time units 
time <- ncvar_get(ncin,"time")
time

tunits <- ncatt_get(ncin,"time","units")
nt <- dim(time)
#How many time units there are 
nt
#Days since...
tunits 

#What kind of calendar? 
ncin$dim$time$calendar
#Gregorian

#Get the the variable (tmp) and its attributes, and verify the size of the array.
# get temperature

tmp <- ncvar_get(ncin,"sst")
st <- dim(tmp)

tmp_array <- ncvar_get(ncin, "sst")


#dlname <- ncatt_get(ncin, "sst","long_name")
#dunits <- ncatt_get(ncin,"sst","units")
#fillvalue <- ncatt_get(ncin, "sst","_FillValue")
#dim(tmp_array)

# convert time -- split the time units string into fields
#tustr <- strsplit(tunits$value, " ")
#tdstr <- strsplit(unlist(tustr)[3], "-")
#tmonth <- as.integer(unlist(tdstr)[2])
#tday <- as.integer(unlist(tdstr)[3])
#tyear <- as.integer(unlist(tdstr)[1])
#new_time <- chron(time,origin=c(tmonth, tday, tyear))

#new_time

# replace netCDF fill values with NA's - no fill values found 
#tmp_array[tmp_array==fillvalue$value] <- NA

#How much of the data are land points? 
#length(na.omit(as.vector(tmp_array[,,1])))

# get a single slice or layer (January)
#Changed to temp1[m]
#m <- 1
#tmp_slice <- tmp[,,m]
#dim(tmp_slice)

#Using the time slice above to look at January temps 
#image(lon,lat,tmp_slice, col=rev(brewer.pal(10,"RdBu")))

#All temps 
# levelplot of the slice - doesn't look right 
#grid <- expand.grid(lon=lon, lat = lat)
#cutpts <- c(-20,-10,-5,-2,0,2,5,10,20,30)
#levelplot(tmp_slice ~ lon * lat, data=grid, at=cutpts, cuts=11, pretty=T, 
          #col.regions=(rev(brewer.pal(10,"RdBu"))))

ncin$dim$time$units

library(ncdf4.helpers)
library(PCICt)
library(dplyr)

#This use of non-standard calendars messes up the calculations of as.Date, 
#so you must use a more specialized function to read in the time variable of the
#climate model output and map it to standard calendar dates. 
#You can do this with the nc.get.time.series function from the ncdf4.helpers package
#(you should also load the PCICt package to use this function):

#This sets the date to a standard: YYYYMMDD and time as well
sst_time <- nc.get.time.series(ncin, v = "sst",
                               time.dim.name = "time")
sst_time[c(1:3, length(sst_time) - 2:0)]

#Getting sst data 
sst <- ncvar_get(ncin, "sst")

#This variable is in a 3-dimensional array, with dimensions ordered as first longitude, then latitude, then time:
#Checking on SST data 
dim(sst)


length(sst_time)
summary(sst_time)

#Selecting a specific time point and place 
#Answers are in celcius? CHECK
#Looking at specific lat/long in September 2011 
#lon_index <- which.min(abs(lon - -5.72))
#lat_index <- which.min(abs(lat - 50.06))
#time_index <- which(format(sst_time, "%Y-%m-%d") == "2000-09-16")
#sst[lon_index, lat_index, time_index]

#Selecting a specific lat/long section 
#lon_index <- which(ncin$dim$lon$vals > -11 | ncin$dim$lon$vals < 3)
#lat_index <-  which(ncin$dim$lat$vals > 49 & ncin$dim$lat$vals < 60.9)

#ts.bounds <- nc.make.time.bounds(ts, unit="month")


#sst[lon_index, lat_index, time_index]

#Selecting only one point 
#lon_index <- which.min(abs(lon - 1.75))
#lat_index <- which.min(abs(lat - 52.48))
#time_index <- as.PCICt(c("1913-01-01", "2016-01-01"), cal="360")
#sst[lon_index, lat_index, time_index]


#Picking out specific of data - this is one lat/long from 1870-2017
#Note that this code uses format and as.Date to convert the PCICt object 
#to a date object, to allow use of a date axis when plotting with ggplot2.
lon_index <- which.min(abs(lon - - 6.04))
lat_index <- which.min(abs(lat - 52.98))

sst <- nc.get.var.subset.by.axes(ncin, "sst",
                                 axis.indices = list(X = lon_index,
                                                     Y = lat_index))

data_frame(time = sst_time,
           sst = as.vector(sst)) %>% 
  mutate(time = as.Date(format(sst_time, "%Y-%m-%d"))) %>%
#Write csv is for extracting csv data and compare each site 
#write.csv(file = "Wicklow_temp.csv")
#read csv 
#Ireland_Wicklow <- read.csv("Wicklow_temp.csv")


#Add the below for graphing 
  ggplot(aes(x = time, y = sst)) + 
  geom_line() + 
  xlab("Date") + ylab("Temperature (C)") + 
  ggtitle("Daily measured sea-surface-temperature, 1870 - 2017",
          subtitle = "At model grid point nearest Dover, UK") + 
  theme_classic()
```

##Pulling together the temperature data for Ireland and the UK
```{r setup, include=FALSE}

#Pulling together all of the data 

#Ireland data binding and cleaning 
Ireland <- bind_cols(Ireland_BrowHead, Ireland_Cleggan, Ireland_Wicklow)
  
Ireland <- Ireland %>%
    dplyr::rename(sst_Browhead_IR = sst) %>%
    dplyr::rename(sst_Cleggan_IR = sst1) %>%
    dplyr::rename(sst_Wicklow_IR = sst2) 

Ireland <- Ireland %>%
select(time, sst_Browhead_IR, sst_Cleggan_IR, sst_Wicklow_IR)
  

#UK data binding and cleaning 
UK <- bind_cols(UK_Ballyhillin, UK_Barrow, UK_Dover, UK_GoreCliff, UK_Holyhead, UK_JohnOGroats, 
                UK_LandsEnd, UK_Lindisfarne, UK_Lowerstoft, UK_Mull, UK_Shetland)

  
  UK <- UK %>%
    dplyr::rename(sst_Ballyhillin_UK = sst) %>%
    dplyr::rename(sst_Barrow_UK = sst1) %>%
    dplyr::rename(sst_Dover_UK = sst2) %>%
    dplyr::rename(sst_GoreCliff_UK = sst3) %>%
    dplyr::rename(sst_Holyhead_UK = sst4) %>%
    dplyr::rename(sst_JohnOGroats_UK = sst5) %>%
    dplyr::rename(sst_LandsEnd_UK = sst6) %>%
    dplyr::rename(sst_Lindisfarne_UK = sst7) %>%
    dplyr::rename(sst_Lowerstoft_UK = sst8) %>%
    dplyr::rename(sst_Mull_UK = sst9) %>%
    dplyr::rename(sst_Shetland_UK = sst10) 
  
  UK <- UK %>%
    select(time, sst_Ballyhillin_UK, sst_Barrow_UK, sst_Dover_UK, sst_GoreCliff_UK, sst_Holyhead_UK,
           sst_JohnOGroats_UK, sst_LandsEnd_UK, sst_Lindisfarne_UK, sst_Lowerstoft_UK, sst_Mull_UK, 
           sst_Shetland_UK)
  
#Binding the UK and Ireland 
UK_Ireland <- bind_cols(UK, Ireland)
#Removing the second time variable (from the Ireland datsset) as it is the same as the UK time 
#column and a duplicate is not needed 

UK_Ireland$time1 <- NULL

write.csv(UK_Ireland, file = "UK_Ireland_SST.csv")
#Have added the mean

UK_Ireland_SST <- read.csv("UK_Ireland_SST.csv")


#Plotting the temperature data 

plot(UK_Ireland_SST)
#group = 1 tells R that the points should be connected for one group 
#group = g no lines drawn
#group = 2 lines are drawn for two seperate groups 

#This code plots Ireland 
ggplot() + 
  geom_line(data = Ireland_BrowHead, aes(x = time, y = sst, group = 1)) + 
  geom_line(data = Ireland_Cleggan, aes(x = time, y = sst, group = 1)) + 
  geom_line(data = Ireland_Wicklow, aes(x = time, y = sst, group = 1)) + 
  labs(y = "SST",
        x = "time") +
  ggtitle("Daily measured sea-surface-temperature, 1870 - 2017",   
          subtitle = "At model grid point nearest several Irish locations") + 
  theme_classic()


#Plots of the UK and Ireland
ggplot() + 
  geom_line(data = UK_Ballyhillin, aes(x = time, y = sst, group = 1)) + 
  geom_line(data = UK_Barrow, aes(x = time, y = sst, group = 1)) + 
  geom_line(data = UK_Dover, aes(x = time, y = sst, group = 1)) + 
  geom_line(data = UK_GoreCliff, aes(x = time, y = sst, group = 1)) + 
  geom_line(data = UK_Holyhead, aes(x = time, y = sst, group = 1)) + 
  geom_line(data = UK_JohnOGroats, aes(x = time, y = sst, group = 1)) + 
  geom_line(data = UK_LandsEnd, aes(x = time, y = sst, group = 1)) + 
  geom_line(data = UK_Lindisfarne, aes(x = time, y = sst, group = 1)) + 
  geom_line(data = UK_Lowerstoft, aes(x = time, y = sst, group = 1)) + 
  geom_line(data = UK_Mull, aes(x = time, y = sst, group = 1)) + 
  geom_line(data = UK_Shetland, aes(x = time, y = sst, group = 1)) +
  geom_line(data = Ireland_BrowHead, aes(x = time, y = sst, group = 1)) + 
  geom_line(data = Ireland_Cleggan, aes(x = time, y = sst, group = 1)) + 
  geom_line(data = Ireland_Wicklow, aes(x = time, y = sst, group = 1)) + 
  geom_smooth(span = 0.3) +
  labs(y = "SST",
       x = "time") +
  ggtitle("Daily measured sea-surface-temperature, 1870 - 2017",   
          subtitle = "At model grid points nearest to several UK and Ireland locations") + 
  theme_classic() 


#Want to plot the mean SST
UK_mean_SST <- UK_Ireland_SST %>%
  select(time, UK_mean) 

#Plotting UK mean temperature 
bb1 <- ggplot() + 
  geom_point(data = UK_mean_SST, aes(x = time, y = UK_mean, group = 1)) + 
  geom_smooth(se = FALSE) +
  scale_colour_gradient(low='yellow', high='#de2d26') +
  labs(y = "SST",
       x = "time") +
  ggtitle("Daily measured sea-surface-temperature, 1870 - 2017",   
          subtitle = "UK mean") + 
  theme_classic()

#Trying to add a trend line 
#bb1<- bb1+ geom_smooth(span = 0.5) + 
  #scale_colour_gradient(low='yellow', high='#de2d26') 

bb1
```


##Extracting yearly max SST from the 14 UK and Ireland sites 
```{r setup, include=FALSE}
#Extracting yearly max SST data 

#Splitting SST into day, month, year 
df <- data.frame(date = UK_mean_SST$Date,
                 year = as.numeric(format(UK_mean_SST$Date, format = "%Y")),
                 month = as.numeric(format(UK_mean_SST$Date, format = "%m")),
                 day = as.numeric(format(UK_mean_SST$Date, format = "%d")))

#Binding SST and dates/months/years and removing the extra 'date' 
SST_day_month_year <- bind_cols(df, UK_mean_SST) 
SST_day_month_year$Date <- NULL

#Average SST for each of the months from 1912 - 2016
aggregate(SST_day_month_year$UK_mean, by = list(SST_day_month_year$year), max)
SST_yearly_max <- aggregate(UK_mean ~ year, data = SST_day_month_year, max)

#Renaming the max_year temp and also selcting from 1913:2015 only (as my data starts
#1912 and ends halfway through 2016)
SST_yearly_max <- SST_yearly_max %>%
  dplyr::rename(year_max = UK_mean) %>%
  filter(year %in% c(1913:2015))

ggplot(data = SST_yearly_max, aes(x = year, y = year_max)) +
  geom_line() + 
  xlab("Year") + ylab("Maximum recorded SST ("~degree~"C)") + 
  ggtitle("Yearly maximum UK sea-surface-temperature (SST) (1913 - 2015)",
          subtitle = "Average taken from 14 sites closest to chosen model grid points") + 
  theme_classic()


##############################################################################################
#Extracting monthly maximum.... 
#Need to go back to the original dataset as the above uses the mean 

#CAN'T GET THIS TO WORK 
#UK_Ireland_SST <- UK_Ireland_SST %>%
  #dplyr::rename(Date = time)

#UK_Ireland_SST <- mutate(UK_Ireland_SST, Date = ymd(Date))

#ef <- data.frame(date = UK_Ireland_SST$Date,
                 #year = as.numeric(format(UK_Ireland_SST$Date, format = "%Y")),
                 #month = as.numeric(format(UK_Ireland_SST$Date, format = "%m")),
                 #day = as.numeric(format(UK_Ireland_SST$Date, format = "%d")))


#max(UK_Ireland_SST_dates[1,5:19])
#UK_Ireland_SST$max <- apply(UK_Ireland_SST[,5:19], max)
#head(UK_Ireland_SST, 10)


UK_Ireland_SST_max <- read.csv("UK_Ireland_SST.csv")

```

#Looking at strandings and SST 

```{r setup, include=FALSE}
#Splitting species data into monthly/quarterly time slots 

#Using the cleaneddata dataset 
library(dplyr)
library(ggplot2)
library(lubridate)


cleaneddata <- read.csv("cleandatesnames.csv")
sapply(cleaneddata, class)

#Having problems with the data format - need dates to be 'date' class
cleaneddata <- mutate(cleaneddata, Date = dmy(Date))
#Trying this instead - works by changing all 2000s to 1900s
#cleaneddata <- mutate(cleaneddata, Date = format(as.Date(Date, "%d-%b-%y"), "19%y-%m-%d"))

#write.csv(cleaneddata, file = "cleandatesnames.csv")

#Now Date should be 'Date' class 

#Selecting the chosen area (if required)
M_novaengliae <- cleaneddata %>% 
  filter(Name.Current.Sci == "Megaptera novaeangliae") 

sapply(M_novaengliae, class)
#Use a variation of the following code if you want to select a specfic window 
#nw_window <- nw_scotland_strandings %>% 
  #filter(Year %in% c(1989:2000))

#Shows this datset by year but I need it by month.....
#ggplot(nw_window, aes(x = Date)) + 
  #stat_count(width = 0.5) 

mn_window <- M_novaengliae %>%
  filter(Year %in% c(1913:2015))

library(zoo)
library(dplyr)

#Splitting my data into strandings per month 
mn_window$monthYear <- as.Date(as.yearmon(mn_window$Date))
mn_window$quarterYear <- as.Date(as.yearqtr(mn_window$Date))
head(mn_window)
#Strandings gathered into months 
mn_monthly <- head(mn_window %>% group_by(monthYear) %>% dplyr::summarise(n = n()), 21)
#Strandings per quarter 
mn_quarter <- head(mn_window %>% group_by(quarterYear) %>% dplyr::summarise(n = n()), 20)

#Grouping by week number - not really needed - not running properly
#pp_window$week <- as.Date("1913-01-01")+7*trunc((pp_window$joinTimestamp / 1000)/(3600*24*7))

#Line plot of strandings per month (grouped)
#Have changed monthYear to Date 
ggplot(data = la_monthly, aes(x = monthYear, y = n, group=1)) +
  geom_line()

ggplot(ba_monthly, aes(n)) +
  geom_histogram(binwidth = 0.5)
  
```

###Plotting mean monthly temperatures 
```{r setup, include=FALSE}

#Plotting species against temperature 
#Example here is Phocoena monthly strandings against STT monthly temps (UK averaged 
#from 14 locations)

#Monthly stranding data is from the above code 
#SST is from the HadISST_Uk.R code 

UK_Ireland_SST <- read.csv("UK_Ireland_SST.csv")

#Renaming time to Date 
UK_Ireland_SST <- UK_Ireland_SST %>%
  dplyr::rename(Date = time)

#Selecting only the UK_mean 
UK_mean_SST <- UK_Ireland_SST %>%
  select(Date, UK_mean) 

#Selecting only 1913:2015
#Need to make Data a date first 
UK_mean_SST <- mutate(UK_mean_SST, Date = ymd(Date))
sapply(UK_mean_SST, class)

write.csv(UK_mean_SST, file = "UK_mean_SST.csv")

#Filtering out specific dates 
UK_mean_SST <- UK_mean_SST %>%
  filter(Date >= "1912-12-16", Date <= "2016-01-16")


#Plotting UK mean temperature 
#Need to play with the months!
bb1 <- ggplot() + 
  geom_line(data = UK_mean_SST, aes(x = Date, y = UK_mean, group = 1)) + 
  #geom_smooth(se = FALSE) +
  #scale_colour_gradient(low='yellow', high='#de2d26') +
  labs(y = "SST",
       x = "time") +
  ylim(0,17) +
  ggtitle("Daily measured sea-surface-temperature, 1870 - 2017",   
          subtitle = "UK mean") + 
  theme_classic()

bb1




```


#Plotting strandings against temp 
```{r setup, include=FALSE}
#Changing "monthYear" to "Date" in species data 
ba_monthly <- ba_monthly %>%
  dplyr::rename(Date = monthYear)
  
  
#PPhocoena monthly strandings and monthly SST
  ggplot() + 
    geom_line(data = UK_mean_SST, aes(x = Date, y = UK_mean)) + 
    geom_line(data = ba_monthly, aes(x = Date, y = n*10)) +
    scale_y_continuous(sec.axis = sec_axis(~./10, name = "Monthly strandings")) +
    labs(y = "SST ("~degree~"C)",
         x = "Year")

#Any way of binding the two datasets? 
#library(plyr)
#combined <- rbind.fill(UK_mean_SST[c("Date", "UK_mean")], pp_monthly[c("Date", "n")])

#Making the date for temperature the first of the month (for that whole month) 
#UK_mean_01_month <- UK_mean_SST

#Make the SST yearly 
#UK_mean_01_month$monthYear <- as.Date(as.yearmon(UK_mean_01_month$Date))
#UK_mean_01_month$quarterYear <- as.Date(as.yearqtr(UK_mean_01_month$Date))
#head(UK_mean_01_month)

#Line plot of strandings per month (grouped)
#Have changed monthYear to Date 
#ggplot(data = UK_mean_01_month, aes(x = monthYear, y = UK_mean, group=1)) +
  #geom_line()


#library(plyr)
#combined <- rbind.fill(UK_mean_01_month[c("monthYear", "UK_mean")], pp_monthly[c("Date", "n")])
combined <- UK_mean_01_month %>% 
  select(UK_mean, monthYear) %>%
  dplyr::rename(Date = monthYear)
  
#Combining uk_mean temp with species data 
combined <- merge(combined, ba_monthly, all = TRUE, by = c('Date'))
#Replacing NA with 0 
combined <- combined %>%
  mutate(n = ifelse(is.na(n),0, n))

#Phocoena strandings (monthly) vs monthly SST
ggplot(data = combined, aes(x = UK_mean, y = n)) +
  geom_point(size = 0.5) +
  labs(x = "Sea Surface Temperature ("~degree~"(C))", y = "Monthly Balaenoptera acutorostrata strandings") +
  #geom_text(aes(label = Date), size = 3, vjust = -0.5) +
  geom_smooth(method = lm, se=FALSE, colour = "grey70", size =0.7) 
```

##Yearly max temp and yearly strandings 
```{r setup, include=FALSE}

#Want to plot yearly max temp against yearly strandings 
#This is total_count 

#Bind STT_yearly max and totalcount 
SST_total_strandings <- bind_cols(SST_yearly_max, totalcount)
#remove duplicate 'Year' 
SST_total_strandings$Year <- NULL
#rename freq to strandings 
SST_total_strandings <- SST_total_strandings %>%
  dplyr::rename(strandings = freq)

#write.csv(SST_total_strandings, file = "SST_strandings_UK.csv")

#Yearly strandings total and yearly max temps 
f1 <- ggplot(data = SST_total_strandings, aes(x = year_max, y = strandings)) +
  geom_point(size = 0.5) +
  geom_text(aes(label = Year), size = 3, vjust = -0.5) +
  labs(x = "Yearly max. sea surface temperature ("~degree~"C)", y = "Yearly total strandings") +
  geom_smooth(method = lm, se=FALSE, colour = "grey70", size =0.7) 

f1

```

```{r setup, include=FALSE}
#Total number of species per year, and maximum temp 
#Using total_speciesbyyear 

#Binding total strandings, total species per year and max years SST 
SST_total_strandings <- bind_cols(total_speciesbyyear, SST_total_strandings)
#Renaming strandings to strandings_total
SST_total_strandings <- SST_total_strandings %>%
  dplyr::rename(strandings_total = strandings)

#Selecting the specific columns
SST_total_strandings <- SST_total_strandings %>%
  select(Year, year_max, species, strandings_total)

#Plotting yearly total species against max yearly SST
ggplot(data = SST_total_strandings, aes(x = year_max, y = species)) +
  geom_point(size = 0.5) +
  geom_text(aes(label = Year), size = 3, vjust = -0.5) +
  labs(x = "Yearly max. sea surface temperature ("~degree~"C)", y = "Total species recorded") +
  geom_smooth(method = lm, se=FALSE, colour = "grey70", size =0.7) 

#Data analysis - total species and yearly max SST
modelSST <- lm(log(species) ~ log(year_max), data = SST_total_strandings)
summary(modelSST)
```


##Monthly max temp and monthly strandings
###Extracting monthly strandings 
```{r setup, include=FALSE}
#Monthly species strandings (only have yearly at the moment)
library(dplyr)
library(plyr)
library(lubridate)
library(zoo)
library(ggplot2)

cleaneddata <- read.csv("cleandatesnames.csv")
#I want all species 

species_window <- cleaneddata %>%
  filter(Year %in% c(1913:2015))

sapply(species_window, class)
#Be cautious of the format: dmy, ymd
species_window <- mutate(species_window, Date = dmy(Date))

sapply(species_window, class)

#Splitting my data into strandings per month 
species_window$monthYear <- as.Date(as.yearmon(species_window$Date))
species_window$quarterYear <- as.Date(as.yearqtr(species_window$Date))
head(species_window)

#Strandings gathered into months 
species_monthly <- head(species_window %>% group_by(monthYear) %>% dplyr::summarise(n = n()), 17402)
#Strandings per quarter 
species_quarter <- head(species_window %>% group_by(quarterYear) %>% dplyr::summarise(n = n()), 1133)

#Grouping by week number - not really needed - not running properly
#pp_window$week <- as.Date("1913-01-01")+7*trunc((pp_window$joinTimestamp / 1000)/(3600*24*7))

#Line plot of strandings per month (grouped)
#Have changed monthYear to Date 
species_monthly <- species_monthly %>%
  dplyr::rename(Date = monthYear)

#Monthly strandings 
ggplot(data = species_monthly, aes(x = Date, y = log(n), group=1)) +
  geom_line() +
  labs(y = "Monthly strandings")

ggplot(species_monthly, aes(n)) +
  geom_histogram(binwidth = 0.5)

#After 1970 only
#Can play around with the dates 
species_monthly_1970 <- species_monthly %>%
  filter(Date >= "1970-01-01", Date <= "2016-01-01")
#Plot
ggplot(data = species_monthly_1970, aes(x = Date, y = log(n), group=1)) +
  geom_line() +
  labs(y = "Monthly strandings since 1970") 

#Quarterley strandings - split into Jan, April, July and October 
#Firstly making date 'Date' 
species_quarter <- species_quarter %>%
  dplyr::rename(Date = quarterYear)

ggplot(data = species_quarter, aes(x = Date, y = log(n), group=1)) +
  geom_line() +
  labs(y = "Quarterly strandings")

#After 1970 only
#Can play around with the dates 
species_quarter_1970 <- species_quarter %>%
  filter(Date >= "1970-01-01", Date <= "2016-01-01")

#Plot
ggplot(data = species_quarter_1970, aes(x = Date, y = log(n), group=1)) +
  geom_line() +
  labs(y = "Quarterly strandings since 1970") 


#Monthly mean SST against monthly strandings 
ggplot() +
  geom_line(data = species_monthly, aes(x = Date, y = n/5)) +
  geom_line(data = UK_mean_SST, aes(x = Date, y = UK_mean)) +
  scale_y_continuous(sec.axis = sec_axis(~.*5, name = "Monthly stranded species")) +
  labs(x = "Date", y = "Monthly mean SST ("~degree~"C)")
```


###Playing around with monthly data and different time scales, with different species 
```{r setup, include=FALSE}
#UK monthly max 
#Change time to date 

UK_Ireland_SST_max <- read.csv("UK_Ireland_SST.csv")
UK_Ireland_SST_max <- UK_Ireland_SST_max %>%
  dplyr::rename(Date = time)

#Specific columns 
UK_monthly_SST <- UK_Ireland_SST_max %>%
  select(Date, Monthly_max)

sapply(UK_monthly_SST, class)


#Filter out the dates 
UK_monthly_SST <- mutate(UK_monthly_SST, Date = ymd(Date))
  UK_monthly_SST <- UK_monthly_SST %>%
           filter(Date >= "1912-12-16", Date <= "2016-01-16")

  
#Monthly max temp and monthly strandings 
ggplot() +
  geom_line(data = species_monthly_1970, aes(x = Date, y = n/10)) +
  geom_line(data = UK_monthly_SST, aes(x = Date, y = Monthly_max)) +
  scale_y_continuous(sec.axis = sec_axis(~.*10, name = "Monthly stranded species")) +
  labs(x = "Date", y = "Monthly maximum SST ("~degree~"C)")

#Binding monthly max temp and monthly strandings 
#Not happy with this as this is only one temp from one location which happens to have the highest 
#value :( 
#Also changing dates to 01 of the month 

#Making the date for temperature the first of the month (for that whole month) 
UK_01_month_max <- UK_monthly_SST

UK_01_month_max$monthYear <- as.Date(as.yearmon(UK_01_month_max$Date))
UK_01_month_max$quarterYear <- as.Date(as.yearqtr(UK_01_month_max$Date))
head(UK_01_month_max)

#Selecting specific rows 
UK_01_month_max <- UK_01_month_max %>% 
  select(Monthly_max, monthYear) %>%
  dplyr::rename(Date = monthYear)

#Merging the two datasets 
monthly_max_species <- merge(UK_01_month_max, species_monthly, all = TRUE, by = c('Date'))
#Replacing NA with 0 
monthly_max_species <- monthly_max_species %>%
  mutate(n = ifelse(is.na(n),0, n))

monthly_max_species <- monthly_max_species %>%
  dplyr::rename(strandings = n)

#Plot of monthly max temp from one single point - yuck and all monthly strandings 
ggplot(data = monthly_max_species, aes(x = Monthly_max, y = strandings)) +
  geom_point(size = 0.5) +
  labs(x = "Monthly max sea surface temperature ("~degree~"(C))", y = "Monthly strandings") +
  #geom_text(aes(label = Date), size = 3, vjust = -0.5) +
  geom_smooth(method = lm, se=FALSE, colour = "grey70", size =0.7) 

#This doesn't work because of the NAs in my dataset, I think 
model_all_species <- lm(log(strandings) ~ log(Monthly_max), data = monthly_max_species)
summary(modelSST)
```


###Remove Phocoena phocoena 







###Plotting specific species against monthly max SST 
```{r setup, include=FALSE}

#Plotting specific species against monthly max SST 
#Merging the two datasets 
#Have to get Date into species data too 

mn_monthly <- mn_monthly %>%
  dplyr::rename(Date = monthYear)

monthly_m_novaengliae <- merge(UK_01_month_max, mn_monthly, all = TRUE, by = c('Date'))
#Replacing NA with 0 
monthly_m_novaengliae <- monthly_m_novaengliae %>%
  mutate(n = ifelse(is.na(n),0, n))

#Plot species and max monthly temp 
ggplot(data = monthly_m_novaengliae, aes(x = Monthly_max, y = n)) +
  geom_point(size = 0.5) +
  labs(x = "Monthly max sea surface temperature ("~degree~"(C))", y = "Monthly Megaptera novaeangliae strandings") +
  #geom_text(aes(label = Date), size = 3, vjust = -0.5) +
  geom_smooth(method = lm, se=FALSE, colour = "grey70", size =0.7) 

```


To do: 
Strandings and monthly STT max/mean without porpoises 


This is as far as I've got with mapping/sense checking the data on a map
Need to extract all of the UK data to get this working correctly 
```{r setup, include=FALSE}
#Modelling temperature for a single day - need to change the code as this currently 
#gives a global picture 
library(ggmap)
library(viridis)
install.packages("weathermetrics")
library(weathermetrics)

#This code maps the temp data for the whole planet 
time_index <- which(format(sst_time, "%Y-%m-%d") == "2016-07-16")
sst <- nc.get.var.subset.by.axes(ncin, "sst",
                                 axis.indices = list(T = time_index))
expand.grid(lon, lat) %>%
  rename(lon = Var1, lat = Var2) %>%
  mutate(lon = ifelse(lon > 180, -(360 - lon), lon),
         sst = as.vector(sst)) %>% 
  mutate(sst = convert_temperature(sst, "c", "k")) %>%
  ggplot() + 
  geom_point(aes(x = lon, y = lat, color = sst),
             size = 0.8) + 
  borders("world", colour="black", fill=NA) + 
  scale_color_viridis(name = "Temperature (C)") + 
  theme_void() + 
  coord_quickmap() + 
  ggtitle("Measured temperature on 2016-07-16",
          subtitle = "HadISST1") 


ts <- as.PCICt(c("1913-01-01", "2016-01-01"), cal="360")
ts.bounds <- nc.make.time.bounds(ts, unit="month")

ncin

```











```
## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
