---
title: "13-extras-species-GAMs"
author: "Ellen Coombs"
date: "11/04/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Extras
### Running generalised additive models (GAMs) for each species 

This RMarkdown document shows the code for running some additional analysis should users want to run GAMs for each individual species, using the same covariates as used in `12-generalised-additive-models.Rmd`. It is not part of the main analysis shown in the paper. 

## Load packages
```{r}
library(dplyr)
library(mgcv) #for GAMs
library(broom) #for pulling all of the code together 
```

## Load data 
```{r}
all_strandings <- read.csv("all_strandings.csv")
```

## Sort individual species data 
This code is required to split out data on each of the individual species. It keeps the original covariate data (i.e., storms, NAO index, sea surface temperature and geomagnetic readings used in the orignal 'all_strandings' GAM in `12-generalised-additive-models.Rmd`)

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
