---
title: "02-cleaning-and-adding-irish-data"
author: "Ellen Coombs"
date: "10/01/2018"
output: html_document
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = FALSE)
```

## Cleaning and importing Irish (IWDG) stranding data 

This script requires the output from `01-clean-up-data-for-analysis.Rmd`. 
The first part of this script imports and cleans the Irish dataset. The second part of this script binds the Irish dataset with the NHM + CSIP dataset (called `cleandatesnames.csv`)

The Irish Whale and Dolphin Group data need to be added here and can be accessed from: 
[Irish Whale and Dolphin Group stranding data] (http://www.iwdg.ie/) 


## Irish data cleaning and binding 

## Load libraries/packages 
```{r}
library(dplyr) #for data manipulation 
library(lubridate) #deals with date format 

```

## Read in clean NHM and CSIP dataset 
This is using the dataset created using `01-clean-up-data-for-analysis.Rmd`

## Read in raw IWDG data 

```{r}

#Irish data 
#Irish data downloaded and saved as "Irish_strandings_raw.csv"
#Cleaning dates and names to be the same as the "cleandatesnames")
#Removing duplicates that appear in both the NHM and IWDG datasets

#Check species names
levels(cleaneddata$Name.Current.Sci)

Irish_data <- read.csv("Irish_strandings_raw.csv")
Irish_data$X. <- NULL

```

## Cleaning the IWDG data 
### Variable names + species names 
This code removes all non-cetaceans from the dataset (e.g. leatherback turtles). It also cleans up unknowns and other name issues. It then filters the dataset to the same timespan (1913-2015) as the NHM + CSIP data. Variables are also given the same names as the NHM + CSIP dataset, and finally, the same variables are selected as were selected from the NHM + CSIP data 

```{r}

#Rename columns so that I can bind the datasets
Irish_data <- Irish_data %>%
  dplyr::rename(Date = Event.Date)

#Remove non cetaceans 
Irish_data <- Irish_data %>% 
  filter(!(Species %in% c("leatherback turtle", "basking shark", "loggerhead turtle", "Kemp's ridley turtle")))

#Double check we just have cetaceans 
levels(Irish_data$Species)

#Unknowns 
#These are listed as "whale species" in the Irish data 
#Changing to the same names as the NHM/CSIP
Irish_data$Species<-as.character(Irish_data$Species)
Irish_data$Species[Irish_data$Species %in% "cetacean species"] <- "Unknown" 
Irish_data$Species[Irish_data$Species %in% "whale species"] <- "Unknown" 
Irish_data$Species[Irish_data$Species %in% "large whale species"] <- "Unknown" 
Irish_data$Species[Irish_data$Species %in% "lagenorhynchus species"] <- "Unknown delphinid" 
Irish_data$Species[Irish_data$Species %in% "common or striped dolphin"] <- "Unknown delphinid" 
Irish_data$Species[Irish_data$Species %in% "dolphin species"] <- "Unknown delphinid" 
Irish_data$Species[Irish_data$Species %in% "pilot/false killer whale"] <- "Unknown odontocete" 
Irish_data$Species[Irish_data$Species %in%  "dolphin species possibly harbour porpoise"] <- "Unknown"
Irish_data$Species[Irish_data$Species %in%  "sei fin or blue whale"] <- "Unknown mysticete" 
Irish_data$Species[Irish_data$Species %in%  "beaked whale species"] <- "Unknown odontocete"

#Filter out the years (1913:2015)
Irish_data <- Irish_data %>%
  filter(Year %in% c(1913:2015))

#Cleaning Scientific names 
Irish_data$Name.Current.Sci<-as.character(Irish_data$Name.Current.Sci)
Irish_data$Name.Current.Sci[Irish_data$Name.Current.Sci %in% "cetacean species"] <- "Unknown" 
Irish_data$Name.Current.Sci[Irish_data$Name.Current.Sci %in% "whale species"] <- "Unknown" 
Irish_data$Name.Current.Sci[Irish_data$Name.Current.Sci %in% "large whale species"] <- "Unknown" 
Irish_data$Name.Current.Sci[Irish_data$Name.Current.Sci %in% "lagenorhynchus species"] <- "Unknown delphinid" 
Irish_data$Name.Current.Sci[Irish_data$Name.Current.Sci %in% "common or striped dolphin"] <- "Unknown delphinid" 
Irish_data$Name.Current.Sci[Irish_data$Name.Current.Sci %in% "dolphin species"] <- "Unknown delphinid" 
Irish_data$Name.Current.Sci[Irish_data$Name.Current.Sci %in% "pilot/false killer whale"] <- "Unknown odontocete" 
Irish_data$Name.Current.Sci[Irish_data$Name.Current.Sci %in%  "dolphin species possibly harbour porpoise"] <- "Unknown"
Irish_data$Name.Current.Sci[Irish_data$Name.Current.Sci %in%  "sei fin or blue whale"] <- "Unknown mysticete" 
Irish_data$Name.Current.Sci[Irish_data$Name.Current.Sci %in%  "beaked whale species"] <- "Unknown odontocete"
Irish_data$Name.Current.Sci[Irish_data$Name.Current.Sci %in%  "dolphin species possibly Phocoena phocoena"] <- "Unknown"
Irish_data$Name.Current.Sci[Irish_data$Name.Current.Sci %in%  "common or Stenella coeruleoalba"] <- "Unknown delphinid"
Irish_data$Name.Current.Sci[Irish_data$Name.Current.Sci %in%  "Pseudorca crassidens "] <- "Pseudorca crassidens"


#Need to rename the variables to be the same as NHM/CSIP for binding
Irish_data <- Irish_data %>% 
  dplyr::rename(Name.Common = Species) %>%
  dplyr::rename(County = County.Region) %>%
  dplyr::rename(Latitude = GPS.Lat) %>%
  dplyr::rename(Longitude = GPS.Long) %>%
  dplyr::rename(S.W.No. = Stranding.ID)

#Select out the columns
iwdg <- Irish_data %>%
  select(Date, Year, Name.Common, Name.Current.Sci, Latitude, Longitude, County, S.W.No.)

```

### Date cleaning and uniformity 
Cleaning the data and making dates the same format at the NHM + CSIP data
```{r}
#Change date format - same as NHM/CSIP
iwdg$Date <- iwdg$Date <- lubridate::dmy(iwdg$Date)
#As factor to make all data the same 
iwdg$Name.Common <- as.factor(iwdg$Name.Common)
iwdg$Name.Current.Sci <- as.factor(iwdg$Name.Current.Sci)

#Double check all 
sapply(cleaneddata, class)
sapply(iwdg, class)

#Cleaneddata Date = factor (may not have to run this - but needed to here)
cleaneddata <- mutate(cleaneddata, Date = dmy(Date))
```

## Adding a data souce column to IWDG dataset 
This code provides a datasource column should it be required for reference
Data columns should now be the same as the cleaneddata (NHM + CSIP dataset)
```{r}
iwdg$Source <- "IWDG"

```

## Merging the two datasets (NHM + CSIP with the IWDG data)
Note: 'cleaneddata' is the NHM + CSIP dataset
```{r}
#Bind cleanedata with Irish data 
UK_and_Irish <- bind_rows(cleaneddata, iwdg)

#Arrange by date 
UK_and_Irish <- UK_and_Irish %>%
  arrange(Date)

arrange(UK_and_Irish, Date)
```


There was the added complication of duplicates in both datasets. This was because data for Ireland was once fed into the NHM dataset - resulting in ~200 duplicates, of which most were labelled "EIRE". The following code: 

- Filters out the EIRE data (these are duplicates) EIRE appears in the NHM data 
- Deletes all data with "N.I" prefix. These are duplicates that are covered by the NHM dataset 
- Renames Irish data in the format seen in the IWGD dataset 
- Removes duplicates seen in both the NHM data and the IWDG datasets 

```{r}
#Be aware that Antrim is also written "Antrim N. Ireland"
#Down = Co.Down 
#Antrim = Co. Antrim 
#Delete all N.I then rename Derry to Co. Derry 
#Londonderry to Co. Derry 
UK_and_Irish <- UK_and_Irish %>% 
  dplyr::filter(UK_and_Irish, !grepl("EIRE", County))

#Remove N.I numbers - this removes most duplicates 
UK_and_Irish <- UK_and_Irish %>% 
  dplyr::filter(!(S.W.No. %in% c("N.I.\\.\\d+")))

UK_and_Irish$County[UK_and_Irish$County %in%  "Down"] <- "Co. Down"
UK_and_Irish$County[UK_and_Irish$County %in%  "Antrim"] <- "Co. Antrim"
UK_and_Irish$County[UK_and_Irish$County %in%  "Londonderry"] <- "Co. Derry"
UK_and_Irish$County[UK_and_Irish$County %in%  "Derry, N.Ireland"] <- "Co. Derry"
UK_and_Irish$County[UK_and_Irish$County %in%  "Down, N.Ireland"] <- "Co. Down"
UK_and_Irish$County[UK_and_Irish$County %in%  "Antrim, N.Ireland"] <- "Co. Antrim"
UK_and_Irish$County[UK_and_Irish$County %in%  "Londonderry"] <- "Co. Derry"
UK_and_Irish$County[UK_and_Irish$County %in%  "Co.Derry, N.Ireland"] <- "Co. Derry"
UK_and_Irish$County[UK_and_Irish$County %in%  "Donegal"] <- "Co. Donegal"
UK_and_Irish$County[UK_and_Irish$County %in%  "Cork"] <- "Co. Cork"

UK_and_Irish$Name.Current.Sci<-as.character(UK_and_Irish$Name.Current.Sci) 
UK_and_Irish$Name.Current.Sci[UK_and_Irish$Name.Current.Sci %in%  "Orcinus orca "] <- "Orcinus orca"
UK_and_Irish$Name.Current.Sci[UK_and_Irish$Name.Current.Sci %in%  "Physeter macrocephalus "] <- "Physeter macrocephalus"
UK_and_Irish$Name.Current.Sci[UK_and_Irish$Name.Current.Sci %in%  "Pseudorca crassidens "] <- "Pseudorca crassidens"

#For duplicates by date, name and county 
UK_and_Irish <- UK_and_Irish[!(duplicated(UK_and_Irish[c("Date","Name.Current.Sci", "County")]) | duplicated(UK_and_Irish[c("Date","Name.Current.Sci", "County")], fromLast = TRUE)), ]

#Dataset is saved as the following csv: 
#write.csv(UK_and_Irish, file = "UK_and_Irish_strandings.csv")
#This is now the final cleaned dataset for the analysis
```


## Quick exploration of what's being seen in the combined final dataset (if desired)
```{r}
#A count per year of each species (with unknowns) - if required 
Strandings_count_IRL_UK <- count(Strandings_IRL_UK, Year, Name.Current.Sci)

#Plotting the above if required 
#Temporal plot by species 
ggplot(data = Strandings_count_IRL_UK, aes(x = Year, y = n, colour= Name.Current.Sci))+
  theme(panel.background = element_blank(), panel.border = element_rect(colour = "grey40", fill = NA)) +
  labs(x = "Year", y = "Count") +
  geom_line() +
  theme_bw() + 
  theme(legend.position="bottom") +
  scale_fill_manual(values=c("deeppink", "steelblue"), guide=FALSE) + 
  facet_wrap(~ Species)   

```

