---
title: "08-correlates-data-geomagentic"
author: "Ellen Coombs"
date: "17/01/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Collecting storms data for the model 
No prerequisites are required for this code to run. This code reads in and cleans geomagnetic data which will form part of the dataset of possible correlates of strandings. 

```{r}
library(dplyr) #for mutate 
library(tidyr) #for unnest 
```

## Import raw geomagnetic data
These data were obtained from the British Geological Survey. 
```{r}
#This is the global geomag data
geomag <- read.csv("Geomagnetic_data.csv")
```


## Formatting the data 
The data were in an odd format so need to be cleaned first 
Raw zip. files are stored in the `raw data` folder 
Description of where the data comes from is provided in the code. 
### Load packages 
```{r}
require(readr)  # for read_csv()
require(purrr)  # for map(), reduce()
require(stringr)
```

### Formatting the data 
This code needs to be run for each of the 3 datafiles: esk.zip, had.zip and ler.zip. The outputs then need to be saved as csv files for further analysis. 

The data: 
Reading in multiple data files from UK oberservatories 
For Lerwick (LER) and Eskdalemuir (ESK) this is 1940 – present
For Hartland (HAD), this folder contains data for Greenwich (GRW 1868 - 1925) and Abinger (ABN 1926 - 1956) both based in and near London before these observatory operations were moved to Hartland (1957 – present) in Devon when the electrification of tramlines and railway disturbances made measurements in London impossible

Each data file contains a year of K-index data. This is a 3-houly scale of local geomagnetic activity, weighted to its locality. 
The scale ranged from 0 (quiet) – 9 (most disturbed)
The file format is: Day Month  Year Day-number Eight 3-hourly K-index values

```{r}

# Function to fix the weirdly formatted magnet data
# Natalie Cooper 2017


fix_magnet_data <- function(filename){
  
  # Read in the magnet data
  raw.data <- read_lines(filename)
  
  # Make list of column names
  # You will need to edit these 
  names.col <- c("Day","Month","Year","Day_number","3_1", "3_2", "3_3", "3_4", "3_5", "3_6", "3_7", "3_8")
  
  # Remove whitespace at the front of the first column
  # Replace spaces with commas
  # Remove excess commas
  # Split into columns using commas as delimiters
  # Outputs a list but with separators in right places
  fixed.data <- raw.data %>%
    str_trim() %>%
    str_replace_all(" ", ",") %>%
    str_replace_all(",,,,", ",") %>%
    str_replace_all(",,,", ",") %>%
    str_replace_all(",,", ",") %>%
    str_split(",")
  
  # Coerce list into a dataframe and name columns
  df1 <- setNames(do.call(rbind.data.frame, fixed.data), names.col)
  
}

# To run the code...

# Extract all the files from a folder
# Due to the weird file extensions you'll need a folder that
# contains _only_ your magnet data
files <- list.files()

# Read in the files
# Tidy them up
# Combine them all
magnet.data.ler <- map_dfr(files, fix_magnet_data)

# This may throw some warning messages about making factors 
# into characters, which isn't an issue at present
# but may need fixing later

write.csv(magnet.data.ler, file ="Geomag_ler.csv")

#The same code should be run for esk.zip and had.zip data files 

```


## Cleaning the geomagnetic data  
Once the above function has been used to clean data from ESK, HAD and LER, then the following cleaning can be done. Reading in the csv files created using the above function. 

This code takes a maximum daily reading (from 8 x 3 hourly recordings of Kp value), and then takes a maximum yearly reading for LER, ESK and HAD. 
```{r}
#Cleaning up the geomag data 

#Firstly the esk.data (Eskdalemuir)
geomag_esk <- read.csv("Geomag_esk.csv")

geomag_esk <- geomag_esk %>%
  select(Day, Month, Year, Mean_daily, Max_daily)

#Take yearly max value from ESK 
#Rename variables 
esk_yearly_max <- aggregate(geomag_esk$Max_daily, by = list(geomag_esk$Year), max)
esk_yearly_max <- esk_yearly_max %>%
  rename(Year = Group.1) %>%
  rename(K_index = x)

#Plot if wanted 
#ggplot(data = esk_yearly_max, aes(x = Year, y = K_index, group = 1)) +
  #geom_line() +
  #labs(y = "Kp value")

#Hartland (HAD)
geomag_had <- read.csv("Geomag_had.csv")
geomag_had <- geomag_had %>%
  select(Day, Month, Year, Mean_daily, Max_daily)

#Take yearly max value from HAD
#Rename variables 
had_yearly_max <- aggregate(geomag_had$Max_daily, by = list(geomag_had$Year), max)
had_yearly_max <- had_yearly_max %>%
  rename(Year = Group.1) %>%
  rename(K_index = x)

#Plot if wanted 
#ggplot(data = had_yearly_max, aes(x = Year, y = K_index, group = 1)) +
  #geom_line() +
  #labs(y = "Kp value")


#Lerwick (LER) 
geomag_ler <- read.csv("Geomag_ler.csv")
geomag_ler <- geomag_ler %>%
  select(Day, Month, Year, Mean_daily, Max_daily)

#Take yearly max value from LER
#Rename variables 
ler_yearly_max <- aggregate(geomag_ler$Max_daily, by = list(geomag_ler$Year), max)
ler_yearly_max <- ler_yearly_max %>%
  rename(Year = Group.1) %>%
  rename(K_index = x)

#Plot if wanted 
#ggplot(data = ler_yearly_max, aes(x = Year, y = K_index, group = 1)) +
  #geom_line() +
  #labs(y = "Kp value")

```


## Combining all 3 datasets 
This code takes the maximum readings from ESK, LER and HAD and filters and combines them so that there is a full dataset of Kp readings from 1913-2015. 

Where there is only one reading for the year (e.g. 1913-1926, was Greenwich (GRW) contained in the HAD datafile), then that yearly max reading is used. If there are multiple readings per year (e.g. ESK and ABN (also contained in the HAD datafile), then a mean max yealry value is taken. Further detials are in the supplementary materials. 

```{r}

#As mentioned 
#For Lerwick (LER) and Eskdalemuir (ESK) this is 1940 – present
#For Hartland (HAD), this folder contains data for Greenwich (GRW 1868 - 1925) 
#and Abinger (ABN 1926 - 1956) both based in and near London before these observatory 
#operations were moved to Hartland (1957 – present) in Devon when the electrification of 
#tramlines and railway disturbances made measurements in London impossible

#Each data file contains a year of K-index data. This is a 3-houly scale of local geomagnetic activity, weighted to its locality. 
#The scale ranged from 0 (quiet) – 9 (most disturbed)


#Combining all the data 
#1913-1940: Had 
#1940-2015: All 

geomag_had1913 <- geomag_had %>%
  filter(Year %in% c(1913:1939))

geom_had1940 <- geomag_had %>%
  filter(Year %in% c(1940:2015))

geom_esk1940 <- geomag_esk %>%
  filter(Year %in% c(1940:2015))

geom_ler1940 <- geomag_ler %>%
  filter(Year %in% c(1940:2015))


#Combining all of the data 
geomag_1940_2015 <- bind_cols(geom_ler1940, geom_esk1940, geom_had1940)

geomag_1940_2015 <- geomag_1940_2015 %>%
  select(Day, Month, Year, Max_daily, Max_daily1, Max_daily2)

#write.csv(geomag_1940_2015, file = "Geomag_1940_2015.csv")


```

## Taking the mean maximum Kp value 
Taking the mean of all the maxes from each of the geomagnetic stations, rather than the 
absolute max 
```{r}

#Need geomag_1940_2015 and take mean max from this + geomag_had1913 as these are the only records for 
#1913 - 1939 (so no mean can be taken)

#Get the mean max values from geomag_1940_2015 
geomag_1940_2015_mean_max <- aggregate(geomag_1940_2015$Mean_max, by = list(geomag_1940_2015$Year), max)
#Rename for binding 
geomag_1940_2015_mean_max<- geomag_1940_2015_mean_max %>%
rename(Year = Group.1) %>%
  rename(Max_K_index = x)

#Bind with geomag_had1913 
All_geom_mean_max <- bind_rows(geomag_1940_2015_mean_max, geomag_1913_max)
All_geom_mean_max <- arrange(All_geom_mean_max, Year)

ggplot(data = All_geom_mean_max, aes(x = Year, y = Max_K_index, group = 1)) +
  geom_line() +
  labs(y = "Max Kp value")

#This is the final dataset which will be fed into the model for Generalised Additive Modelling
write.csv(All_geom_mean_max, file = "Geom_mean_max.csv")

```