---
title: "09-correlates-data-SST"
author: "Ellen Coombs"
date: "10/01/2018"
output: html_document
---



## Collecting sea surface temperature (SST) data for the model 
No prerequisites are required for this code to run. This code reads in and cleans sea surface temperature (SST) data which will form part of the dataset of possible correlates of strandings. 

More details are available in the supplementary material. 
Extracting data from the HadISST (Met Office) dataset was done first. This was carried out for 14 locations around the UK and Ireland (see supplementary material)

These data are from the UK Met Office, they are monthly SST readings. 
## Load packages 
```{r}
library(lattice) #If you want to plot SST maps - visuals 
library(ncdf4) #Reading in netcdf data 
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(ncdf4.helpers)#collection of functions for dealing with NetCDF files opened using ncdf4
library(PCICt) #identifies meaningful correlations to define edges in a weighted network
```

## Load data
The data are downloaded from the Met Office website 
```{r}
#Reading in .nc data 
ncin <- nc_open("HadISST_sst.nc")

#Taking a look at the data 
ncin_data <- ncvar_get(ncin, attributes(ncin$var)$names[1])

```


## Data cleaning and manipulation 
```{r}
 
#This use of non-standard calendars messes up the calculations of as.Date, 
#so you must use a more specialized function to read in the time variable of the
#climate model output and map it to standard calendar dates. 
#You can do this with the nc.get.time.series function from the ncdf4.helpers package
#you should also load the PCICt package to use this function):

#This sets the date to a standard: YYYYMMDD and time as well
sst_time <- nc.get.time.series(ncin, v = "sst",
                               time.dim.name = "time")
sst_time[c(1:3, length(sst_time) - 2:0)]

#Getting sst data 
sst <- ncvar_get(ncin, "sst")

#This variable is in a 3-dimensional array, with dimensions ordered as first longitude, then latitude, then time:
#Checking on SST data - use the following 
dim(sst)
length(sst_time)
summary(sst_time)

```

## Pulling out SST data for multiple locations 
See supplementary material for lat/long of 14 locations around the UK and Ireland. Each location should be put into the below code: long_index and lat_index 

Once this code has been run for each of the 14 sites, a dataframe of SST should be available for each of the sites, e.g. Ireland Wicklow. The HadISST dataset contains readings SST readings all around the UK - other lat/long coordinates can be added to the following, if desired. 

See: [Met Office](https://www.metoffice.gov.uk/hadobs/hadisst/>)

The following lat/longs were used for 14 locations around the UK and Ireland. Each lat/long for each of the 14 sites needs to be run, and then saved as a seperate file. The below code gives and example for County Wicklow ("Ireland_Wicklow")

Use the following lat/longs to pull out SST data for each of the 14 locations. 

---
output: pdf_document
---

| Location        | Latitude      | Longitude            |
| :------------   | :-----------: | -------------------: |
| Ballyhillin, UK	|  55.34        | -7.4                 |
| Second row      | **Cell**      | *Cell*               |
| Third row       | Cell that spans across two columns  ||

Table: Your Caption


```{r}
#Selecting only one location (lat/long)
lon_index <- which.min(abs(lon - 1.75))
lat_index <- which.min(abs(lat - 52.48))
time_index <- as.PCICt(c("1913-01-01", "2016-01-01"), cal="360")
sst[lon_index, lat_index, time_index]


#Picking out specifics of data - this is one lat/long from 1870-2017
#Note that this code uses format and as.Date to convert the PCICt object 
#to a date object, to allow use of a date axis when plotting with ggplot2.
lon_index <- which.min(abs(lon - - 6.04)) 
lat_index <- which.min(abs(lat - 52.98))

sst <- nc.get.var.subset.by.axes(ncin, "sst",
                                 axis.indices = list(X = lon_index,
                                                     Y = lat_index))

data_frame(time = sst_time,
           sst = as.vector(sst)) %>% 
  mutate(time = as.Date(format(sst_time, "%Y-%m-%d")))
  #Write csv is for extracting csv data and compare each site 
#E.g  write.csv(file = "Ireland_Wicklow.csv") 


#To plot if required 
  ggplot(aes(x = time, y = sst)) + 
  geom_line() + 
  xlab("Date") + ylab("Temperature (C)") + 
  ggtitle("Daily measured sea-surface-temperature, 1870 - 2017",
          subtitle = "At model grid point nearest Dover, UK") + 
  theme_classic()

```

## Pulling together all of the data 
This code pulls together the SST data from each of the chosen 14 sites 
```{r}
#Ireland data binding and cleaning 
Ireland <- bind_cols(Ireland_BrowHead, Ireland_Cleggan, Ireland_Wicklow)
#Rename sst1, sst2 etc to the places they represent 
Ireland <- Ireland %>%
    dplyr::rename(sst_Browhead_IR = sst) %>%
    dplyr::rename(sst_Cleggan_IR = sst1) %>%
    dplyr::rename(sst_Wicklow_IR = sst2) 

#Select column specifics 
Ireland <- Ireland %>%
select(time, sst_Browhead_IR, sst_Cleggan_IR, sst_Wicklow_IR)

#UK data binding and cleaning 
UK <- bind_cols(UK_Ballyhillin, UK_Barrow, UK_Dover, UK_GoreCliff, UK_Holyhead, UK_JohnOGroats, 
                UK_LandsEnd, UK_Lindisfarne, UK_Lowerstoft, UK_Mull, UK_Shetland)

#As above - change sst1, sst2 etc to the places they represent 
  UK <- UK %>%
    dplyr::rename(sst_Ballyhillin_UK = sst) %>%
    dplyr::rename(sst_Barrow_UK = sst1) %>%
    dplyr::rename(sst_Dover_UK = sst2) %>%
    dplyr::rename(sst_GoreCliff_UK = sst3) %>%
    dplyr::rename(sst_Holyhead_UK = sst4) %>%
    dplyr::rename(sst_JohnOGroats_UK = sst5) %>%
    dplyr::rename(sst_LandsEnd_UK = sst6) %>%
    dplyr::rename(sst_Lindisfarne_UK = sst7) %>%
    dplyr::rename(sst_Lowerstoft_UK = sst8) %>%
    dplyr::rename(sst_Mull_UK = sst9) %>%
    dplyr::rename(sst_Shetland_UK = sst10) 
  
#Select column specifics 
  UK <- UK %>%
    select(time, sst_Ballyhillin_UK, sst_Barrow_UK, sst_Dover_UK, sst_GoreCliff_UK, sst_Holyhead_UK,
           sst_JohnOGroats_UK, sst_LandsEnd_UK, sst_Lindisfarne_UK, sst_Lowerstoft_UK, sst_Mull_UK, 
           sst_Shetland_UK)
  
#Binding the UK and Ireland 
UK_Ireland <- bind_cols(UK, Ireland)
#Removing the second time variable (from the Ireland dataset)
UK_Ireland$time1 <- NULL

#Renaming the dataset 
UK_Ireland <- UK_Ireland_SST

#Save if required
#write.csv(UK_Ireland_SST, file = "UK_Ireland_SST.csv")

#Plot all temperatures if required 
#Plots of the UK and Ireland
#ggplot() + 
  #geom_line(data = UK_Ballyhillin, aes(x = time, y = sst, group = 1)) + 
  #geom_line(data = UK_Barrow, aes(x = time, y = sst, group = 1)) + 
  #geom_line(data = UK_Dover, aes(x = time, y = sst, group = 1)) + 
  #geom_line(data = UK_GoreCliff, aes(x = time, y = sst, group = 1)) + 
  #geom_line(data = UK_Holyhead, aes(x = time, y = sst, group = 1)) + 
  #geom_line(data = UK_JohnOGroats, aes(x = time, y = sst, group = 1)) + 
  #geom_line(data = UK_LandsEnd, aes(x = time, y = sst, group = 1)) + 
  #geom_line(data = UK_Lindisfarne, aes(x = time, y = sst, group = 1)) + 
  #geom_line(data = UK_Lowerstoft, aes(x = time, y = sst, group = 1)) + 
  #geom_line(data = UK_Mull, aes(x = time, y = sst, group = 1)) + 
  #geom_line(data = UK_Shetland, aes(x = time, y = sst, group = 1)) +
  #geom_line(data = Ireland_BrowHead, aes(x = time, y = sst, group = 1)) + 
  #geom_line(data = Ireland_Cleggan, aes(x = time, y = sst, group = 1)) + 
  #geom_line(data = Ireland_Wicklow, aes(x = time, y = sst, group = 1)) + 
  #geom_smooth(span = 0.3) +
  #labs(y = "SST",
       #x = "time") +
  #ggtitle("Daily measured sea-surface-temperature, 1870 - 2017",   
          #subtitle = "At model grid points nearest to several UK and Ireland locations") + 
  #theme_classic() 

```

## Working with the max yearly SST 
Pulling out the monthly mean from the 14 locations and then selecting the mean yearly max SST
```{r}

#Calculating the monthly mean from all 14 locations 
#Adding the yearly mean, taken by averaging the 14 locations 
UK_Ireland_SST$UK_mean <- rowMeans(subset(UK_Ireland_SST, select = c(sst_Ballyhillin_UK,
                                                                     sst_Barrow_UK, sst_Dover_UK,
                                                                     sst_GoreCliff_UK,
                                                                     sst_Holyhead_UK,
                                                                     sst_JohnOGroats_UK,
                                                                     sst_LandsEnd_UK,
                                                                     sst_Lindisfarne_UK,
                                                                     sst_Lowerstoft_UK,
                                                                     sst_Mull_UK, sst_Shetland_UK,
                                                                     sst_Browhead_IR,
                                                                     sst_Cleggan_IR, 
                                                                     sst_Wicklow_IR)), na.rm=TRUE)

#Pull out the monthly mean SST
UK_mean_SST <- UK_Ireland_SST %>%
  dplyr::select(time, UK_mean) 

#Renaming uk_mean_SST columns 
UK_mean_SST <- UK_mean_SST %>%
  dplyr::rename(Date = time)


#Filter out the columns we want - this is 1913:2015 
#Filtered this way instead of by year (for example) because the dates were messy 
UK_mean_SST <- UK_mean_SST %>% 
  filter(row_number() %in% 517:1752)

#Change date format of new dataset
sapply(UK_mean_SST, class)
UK_mean_SST <- mutate(UK_mean_SST, Date = dmy(Date))


#Splitting SST into day, month, year 
#This code is for breaking down into day, month, year 
df <- data.frame(date = UK_mean_SST$Date,
                 year = as.numeric(format(UK_mean_SST$Date, format = "%Y")),
                 month = as.numeric(format(UK_mean_SST$Date, format = "%m")),
                 day = as.numeric(format(UK_mean_SST$Date, format = "%d")))


#Binding SST and dates/months/years and removing the extra 'date' 
SST_day_month_year <- bind_cols(df, UK_mean_SST) 
SST_day_month_year$Date <- NULL

#Average SST for each of the months from 1912 - 2016
aggregate(SST_day_month_year$UK_mean, by = list(SST_day_month_year$year), max)
SST_yearly_max <- aggregate(UK_mean ~ year, data = SST_day_month_year, max)

#Renaming the max_year temp and also selecting from 1913:2015 only 
SST_yearly_max <- SST_yearly_max %>%
  dplyr::rename(Year_max = UK_mean) %>%
  filter(year %in% c(1913:2015))

#Plot if desired 
#ggplot(data = SST_yearly_max, aes(x = year, y = Year_max)) +
  #geom_line() + 
  #xlab("Year") + ylab("Maximum recorded SST ("~degree~"C)") + 
  #ggtitle("Yearly maximum UK sea-surface-temperature (SST) (1913 - 2015)",
          #subtitle = "Average taken from 14 sites closest to chosen model grid points") + 
  #theme_classic()

#Rename the columns 
SST_yearly_max <- SST_yearly_max %>% 
  rename(Max_SST = Year_max)

#Save for use in dataframe for correlates model 
write.csv(SST_yearly_max, file = "SST_yearly_max.csv")

```


