---
title: "11-correlates-data-fishing-catch"
author: "Ellen Coombs"
date: "07/12/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Collecting North Atlantic Oscillation (NAO) data for the model 

This code is for reading in the fishing catch data. 
Data are from the International Council for the Exploration of the Sea [ICES](www.ices.dk). 

We combined 3 separate fishing datasets into a .csv file: 

ICES Historical Landings: 1903 – 1949 
Historical Nominal Catches: 1950 – 2010 
Official Nominal Catches: 2006 – 2016 

The following code cleans the data for the model. 

##Cleaning the fishing catch data
```{r, echo=TRUE}

Fisheries_data <- read.csv("Landings.csv")

#Country codes for UK and Ireland (inclduing Isle of Man etc.)
Fisheries_data <- Fisheries_data %>%
  filter(Country %in% c("GBE", "GBU", "GBS", "IRL"))

Fisheries_data$X <- NULL
Fisheries_data$X.1 <- NULL
Fisheries_data$X.2 <- NULL
Fisheries_data$X.3 <- NULL 
Fisheries_data$X.4 <-NULL 

#Years selected 
Fisheries_data <- Fisheries_data %>%
  filter(Year %in% c(1913:2015))

#Selecting specific columns
Fisheries_data <- Fisheries_data %>%
  select(Year, Landings)

Fisheries_data$Landings = as.numeric(as.character(Fisheries_data$Landings))
#NAs -> 0 
Fisheries_data[is.na(Fisheries_data)] <- 0

#Aggrgate by year 
Fisheries_data <- aggregate(Fisheries_data$Landings, by=list(Year=Fisheries_data$Year), FUN=sum)

#To save the data
write.csv(Fisheries_data, file = "Fisheries_data.csv")

```

