---
title: "13-extras-species-GAMs"
author: "Ellen Coombs"
date: "11/04/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Extras
### Running generalised additive models (GAMs) for each species 

This RMarkdown document shows the code for running some additional analysis should users want to run GAMs for each individual species, using the same covariates as used in `12-generalised-additive-models.Rmd`. It is not part of the main analysis shown in the paper. 

## Load packages
```{r}
library(dplyr)
library(mgcv) #for GAMs
library(broom) #for pulling all of the code together 
```

## Load data 
```{r}
all_strandings <- read.csv("all_strandings.csv")
```

## Sort individual species data 
This code is required to split out data on each of the individual species. It keeps the original covariate data (i.e., storms, NAO index, sea surface temperature and geomagnetic readings used in the orignal 'all_strandings' GAM in `12-generalised-additive-models.Rmd`)

```{r}
#Species GAMs 
#This code runs the GAMs for each of the indiviual species 

#First, code to pick out each of the seperate species 
#Now we need to take each species from the all_strandings dataset 
#Creating a new data frame to play with
sep_species <- all_strandings

#Make new dataframe by splitting into seperate species (these are tibbles)
lDf <- split(sep_species, sep_species$Species)

#Example of calling up sei whale data 
lDf$`Balaenoptera borealis`

#Splitting the tibble into data frames for each species 
Y <- lapply(seq_along(lDf), function(x) as.data.frame(lDf[[x]])[, 1:8]) 

BA <- Y[[1]] #minke
BB <- Y[[2]] #sei 
BM <- Y[[3]] #blue
BP <- Y[[4]] #fin
DD <- Y[[5]] #common
GM <- Y[[6]] #globicephala
GG <- Y[[7]] #griseus
HA <- Y[[8]] #northern 
KB <- Y[[9]] #pygmy
LA <- Y[[10]] #white sided 
LAl <- Y[[11]] #white beaked 
MN <- Y[[12]] #humpback
MB <- Y[[13]] #Sowerby's
MM <- Y[[14]] #True's 
OO <- Y[[15]] #orca 
PP <- Y[[16]] #harbour 
PM <- Y[[17]] #sperm
PC <- Y[[18]] #false 
SC <- Y[[19]] #striped 
TT <- Y[[20]] #bottlenose
ZC <- Y[[21]] #Cuvier's

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
